Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2023-05-06 Manuscript/Manuscript.Rmd
MS: updated affiliations, changed to knitr to include figures

---
title: Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance
  in zooplankton
author: "*Kei Jokura, Nobuo Ueda, Martin Gühmann, Luis Alfonso Yañez-Guerra, Piotr Słowiński, Kyle C. A. Wedgwood, Gáspár Jékely*"
date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4
  4QD, UK*"
output:
  pdf_document:
    latex_engine: lualatex
    template: default
  html_document:
    theme: lumen
  word_document: default
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
monofont: Times New Roman
documentclass: article
classoption: onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
---

@@

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a three-day-old Platynereis larva. **(B, C)** Five types of neurons (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a three-day-old larva. Ciliated cells labelled with (B) are shown in grey. The INNOS on the ventral side labelled with (C) are shown in purple. Axons and dendrites appear as lines and cell body positions are represented by spheres. **(D)** Expression analysis of the NOS gene (magenta) using in situ HCR. Anterior view of the larva at two-day-old. Co-stained image using acetylated α-tubulin antibody (acTub: green). **(E)** Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green). Views of larvae at three-day-old from the anterior side. Insert: One of the reconstracted INNOS is shown from the anterior side. **(F)** Wiring diagram of the cPRC circuit. Hexagons and arrows represent cell groups and their synaptic connections, respectively. Numbers inside hexagons indicate the number of cells grouped in each node. Arrow line thickness (synapse weight) is equal to the number of synapses.](../figures/Fig1.png){width="100%"}

@@

![**Figure 2. NO produced by UV/violet stimulation to cPRCs.** **(A)** A NO-positive cell visualised by DAF-FM (green) is shown (dashline). The white line indicates the outer frame of the larva. All subsequent images are of two-day-old larvae. **(B)** Intensity changes before (upper) and during 405 nm light stimulation (lower) in a cell visualised by DAF-FM. **(C)** The intensity of DAF-FM changes over time when cPRC (green) and other cells (ctr: gray) are stimulated with UV/violet (405 nm) light. Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean.](../figures/Fig2.png){width="100%"}

@@

![**Figure 3. Larvae with inhibited NOS have a delayed response to UV-avoidance.** **(A)** Top: The domain and Exon/Intron structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Gray shows PAM sequences, red shows stop codons. **(B)** Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) at three-day-old larvae. 0 sec as the starting point. After 10 sec, UV (395 nm) stimulation from the side. **(C)** The temporal changes in the vertical position of the WT and mutant larvae before (dark) and after UV stimulation (side 395 nm) are shown. The starting points of each larval trajectory are set to 0. **(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. **(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins.](../figures/Fig3.png){width="100%"}

@@

![**Figure 4. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR. Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). **(C, D)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub). **(E, F)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E: NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae. **(G, H)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and 2) larvae. **(I)** The domain structure of NIT-GC1. Below shows the mutant NIT-GC1 (ΔNIT-GC1) with deletion of the NIT domain. Grey squares indicate the transmembrane region (TM). **(J)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. **(K-N)** The intensity changes in cGMP fluorescence over time for the four conditions are shown. Gray traces show the optical responses to individual intensities (n \> 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig4.png){width="100%"}

@@

![**Figure 5. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** **(A, B)** Each left figure shows a calcium imaging using GCaMP6s. The respective right figures show immunostaining images of the same positions in the respective left figures, (A) shows RYamide antibodies, (B) shows RGWamide and serotonin (5-HT) antibodies in red and DAPI in cyan. Asterisks indicate the nuclei of the cPRC, respectively, and the nuclei of the cells that were identified from the arrangement of the surrounding cells are indicated by numbers. From the position of the cells, the calcium imaging data identify INNOS in (A) and INRGW and Ser-h1 in (B), respectively. **(C)** Correlation (Pearson's r) map of neuronal activity of the INNOS, INRGW, Ser-h1 and MC neurons. **(D)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light purple) larvae. **(E)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: pink and light pink) larvae. **(F)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in WT (dark blue) and NOSΔ23/Δ23 mutant (light blue) larvae. **(G)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue). **(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae.](../figures/Fig5.png){width="100%"}

@@

![**Figure 6. Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq. The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells. The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene. **(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Fig6.png){width="100%"}

>>>>

---
title: Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance
  in zooplankton
author: 
  - name: "Kei Jokura<sup>1,`*`</sup>, Nobuo Ueda<sup>2</sup>, Martin Gühmann<sup>3</sup>, Luis Alfonso Yañez-Guerra<sup>1</sup>, Piotr Słowiński<sup>1</sup>, Kyle C. A. Wedgewood<sup>1</sup> and Gáspár Jékely<sup>1,4,`*`</sup>"
    affiliation: "<sup>1</sup>Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 <br> <sup>2</sup>Okinawa Institute of Science and Technology, Okinawa, Japan<br> <sup>3</sup>School of Biological Sciences, University of Bristol, Bristol, United Kingdom <br> <sup>4</sup>Centre for Organismal Studies (COS), University of Heidelberg, 69120 Heidelberg, Germany  <br> <sup>*</sup>Correspondence`:` kei.jokura @ exeter.ac.uk, gaspar.jekely @ cos.uni-heidelberg.de"
output:
  html_document:
    theme: lumen
  pdf_document:
    latex_engine: lualatex
    template: default
  word_document: default
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
monofont: Times New Roman
documentclass: article
classoption: onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
---

@@

```{r echo=FALSE,fig.cap="**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** **(A)** SEM image of a three-day-old Platynereis larva. **(B, C)** Five types of neurons (cPRC, INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a three-day-old larva. Ciliated cells labelled with (B) are shown in grey. The INNOS on the ventral side labelled with (C) are shown in purple. Axons and dendrites appear as lines and cell body positions are represented by spheres. **(D)** Expression analysis of the NOS gene (magenta) using in situ HCR. Anterior view of the larva at two-day-old. Co-stained image using acetylated α-tubulin antibody (acTub: green). **(E)** Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green). Views of larvae at three-day-old from the anterior side. Insert: One of the reconstructed INNOS is shown from the anterior side. **(F)** Wiring diagram of the cPRC circuit. Hexagons and arrows represent cell groups and their synaptic connections, respectively. Numbers inside hexagons indicate the number of cells grouped in each node. Arrow line thickness (synapse weight) is equal to the number of synapses."}

knitr::include_graphics(path = "figures/Fig1.png")

@@

```{r echo=FALSE,fig.cap="**Figure 2. NO produced by UV/violet stimulation to cPRCs.** **(A)** A NO-positive cell visualised by DAF-FM (green) is shown (dashline). The white line indicates the outer frame of the larva. All subsequent images are of two-day-old larvae. **(B)** Intensity changes before (upper) and during 405 nm light stimulation (lower) in a cell visualised by DAF-FM. **(C)** The intensity of DAF-FM changes over time when cPRC (green) and other cells (ctr: gray) are stimulated with UV/violet (405 nm) light. Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean."}

knitr::include_graphics(path = "figures/Fig2.png")

@@

```{r echo=FALSE,fig.cap="**Figure 3. Larvae with inhibited NOS have a delayed response to UV-avoidance.** **(A)** Top: The domain and Exon/Intron structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Gray shows PAM sequences, red shows stop codons. **(B)** Overlaid trajectories for WT (n=32) and NOS mutant (NOSΔ11/Δ11, n=26 and NOSΔ23/Δ23, n=47) at three-day-old larvae. 0 sec as the starting point. After 10 sec, UV (395 nm) stimulation from the side. **(C)** The temporal changes in the vertical position of the WT and mutant larvae before (dark) and after UV stimulation (side 395 nm) are shown. The starting points of each larval trajectory are set to 0. **(D)** Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. **(E)** Vertical swimming in larvae treated with NOS inhibitors, L-NAME at three-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins."}

knitr::include_graphics(path = "figures/Fig3.png")

@@

```{r echo=FALSE,fig.cap="**Figure 4. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** **(A, B)** Expression analysis of the NIT-GC1 (C) and NIT-GC2 (D) gene (magenta) using in situ HCR. Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). **(C, D)** Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub). **(E, F)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (E: NOSΔ11/Δ11, F: NOSΔ23/Δ23) (light pink) larvae. **(G, H)** Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in NIT-GC1 (G: NIT1) and NIT-GC2 (H: NIT2) knockdown by morpholino (MO1 and 2) larvae. **(I)** The domain structure of NIT-GC1. Below shows the mutant NIT-GC1 (ΔNIT-GC1) with deletion of the NIT domain. Grey squares indicate the transmembrane region (TM). **(J)** A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. **(K-N)** The intensity changes in cGMP fluorescence over time for the four conditions are shown. Gray traces show the optical responses to individual intensities (n > 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars)."}

knitr::include_graphics(path = "figures/Fig4.png")

@@

```{r echo=FALSE,fig.cap="**Figure 5. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** **(A, B)** Each left figure shows a calcium imaging using GCaMP6s. The respective right figures show immunostaining images of the same positions in the respective left figures, (A) shows RYamide antibodies, (B) shows RGWamide and serotonin (5-HT) antibodies in red and DAPI in cyan. Asterisks indicate the nuclei of the cPRC, respectively, and the nuclei of the cells that were identified from the arrangement of the surrounding cells are indicated by numbers. From the position of the cells, the calcium imaging data identify INNOS in (A) and INRGW and Ser-h1 in (B), respectively. **(C)** Correlation (Pearson's r) map of neuronal activity of the INNOS, INRGW, Ser-h1 and MC neurons. **(D)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in WT (dark purple) and NOSΔ11/Δ11 mutant (light purple) larvae. **(E)** Changes over time in GCaMP6s intensities in INNOS when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: pink and light pink) larvae. **(F)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in WT (dark blue) and NOSΔ23/Δ23 mutant (light blue) larvae. **(G)** Changes over time in GCaMP6s intensities in INRGW when cPRC is stimulated with UV in NIT-GC2 morphants (MO1 and MO2: blue and light blue). **(H, I)** Changes over time in GCaMP6s intensities in Ser-h1 and MC when cPRC is stimulated with UV in WT (Ser-h1 and MC: dark red and brown) and NOSΔ11/Δ11 mutant (Ser-h1 and MC: light red and brown) larvae."}

knitr::include_graphics(path = "figures/Fig5.png")

@@

```{r echo=FALSE,fig.cap="**Figure 6. Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** **(C)** Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq. The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells. The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene. **(D)** Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback."}

knitr::include_graphics(path = "figures/Fig6.png")
